---
title: "Exercise 6"
format:
  html:
    embed-resources: true
    code-fold: true
    warning: false
    message: false
output-file: index.html
---

## Preparation

```{r}
library("readr")
library("sf")
library(dplyr)
library(ggplot2)
library(tmap)
library(terra)
library(lubridate)

wildschwein_BE <- read_delim("Data/wildschwein_BE_2056.csv") |> 
  st_as_sf(coords = c("E", "N"), crs = 2056, remove = FALSE)
```

## Task 1 

```{r}
Feldaufnahmen_Fanel <- read_sf("Data/Feldaufnahmen_Fanel.gpkg")

str(Feldaufnahmen_Fanel)

names(Feldaufnahmen_Fanel)
glimpse(Feldaufnahmen_Fanel)
head(Feldaufnahmen_Fanel)
st_geometry_type(Feldaufnahmen_Fanel)
unique(st_geometry_type(Feldaufnahmen_Fanel))
sapply(Feldaufnahmen_Fanel, class)
st_crs(Feldaufnahmen_Fanel)


ggplot() +
  geom_sf(data = Feldaufnahmen_Fanel, aes(fill = Frucht)) +
  theme_minimal()

tmap_options(basemap.server = 
               "SwissFederalGeoportal.NationalMapColor"  
               )

tmap_mode("view")

tm_shape(Feldaufnahmen_Fanel) + tm_polygons("Frucht")

```

The dataset `Feldaufnahmen_Fanel` was loaded as a simple feature object. It contains **975 features** and **three columns**: `FieldID`, `Frucht`, and `geom`.

The column `FieldID` is numeric and identifies the field. The column `Frucht` is a character column and describes the crop or land-use type, for example **Roggen**, **Wiese**, **Weide**, or **Wald**. Some entries in this column contain missing values.

The geometry column `geom` stores the spatial information. The geometry type of the dataset is **POLYGON**, which means that the data represents areas rather than points or lines.

The coordinate reference system is **CH1903+ / LV95**, corresponding to **EPSG:2056**. This is a projected coordinate reference system used in Switzerland, with coordinates measured in metres.


## Task 2

```{r}

wildschwein_BE <- wildschwein_BE |> 
  filter(
    DatetimeUTC >= "2015-05-01 00:00:00",
    DatetimeUTC <= "2015-06-30 23:59:59"
  )


ggplot() +
  geom_sf(data = Feldaufnahmen_Fanel, aes(fill = Frucht), alpha = 0.5) +
  geom_sf(data = wildschwein_BE, size = 0.3) +
  theme_minimal()


wildschwein_BE_annotated <- st_join(
  wildschwein_BE,
  Feldaufnahmen_Fanel
)

glimpse(wildschwein_BE_annotated)

visited_fruits <- wildschwein_BE_annotated |> 
  count(Frucht, sort = TRUE)


wildschwein_eat_fruits <- wildschwein_BE_annotated |> 
  count(TierName, Frucht, sort = TRUE)

visited_fruits
wildschwein_eat_fruits
```


## Task 3 

```{r}
library(lubridate)
library(forcats)

wildschwein_BE_annotated <- wildschwein_BE_annotated |>
  mutate(
    DatetimeUTC = ymd_hms(DatetimeUTC),
    hour = hour(DatetimeUTC),
    Frucht = if_else(is.na(Frucht), "No field / unknown", Frucht)
  )

crop_hour_counts <- wildschwein_BE_annotated |>
  st_drop_geometry() |>
  count(hour, Frucht, name = "n")


most_common_fruits <- wildschwein_BE_annotated |>
  st_drop_geometry() |>
  count(Frucht, sort = TRUE) |>
  slice_head(n = 5) |>
  pull(Frucht)

crop_hour_percent <- crop_hour_counts |>
  filter(Frucht %in% most_common_fruits) |>
  group_by(hour) |>
  mutate(percent = n / sum(n) * 100) |>
  ungroup()


ggplot(crop_hour_percent, aes(x = hour, y = percent, fill = Frucht)) +
  geom_col(position = "stack") +
  scale_x_continuous(breaks = 0:23) +
  labs(
    title = "Percentage of wild boar samples per crop type and hour",
    x = "Hour of day",
    y = "Percentage of samples",
    fill = "Crop / land-use type"
  ) +
  theme_minimal()


crop_hour_animal <- wildschwein_BE_annotated |>
  st_drop_geometry() |>
  filter(Frucht %in% most_common_fruits) |>
  count(TierName, hour, Frucht, name = "n") |>
  group_by(TierName, hour) |>
  mutate(percent = n / sum(n) * 100) |>
  ungroup()

ggplot(crop_hour_animal, aes(x = hour, y = percent, fill = Frucht)) +
  geom_col(position = "stack") +
  facet_wrap(~ TierName) +
  scale_x_continuous(breaks = seq(0, 23, by = 4)) +
  labs(
    title = "Crop use by hour and individual wild boar",
    x = "Hour of day",
    y = "Percentage of samples",
    fill = "Crop / land-use type"
  ) +
  theme_minimal()
```



## Task 4

```{r}
vegetationshoehe_LFI <- rast("Data/vegetationshoehe_LFI.tif")

vegetationshoehe_LFI
crs(vegetationshoehe_LFI)
res(vegetationshoehe_LFI)
ext(vegetationshoehe_LFI)
nlyr(vegetationshoehe_LFI)

plot(vegetationshoehe_LFI)

tmap_options_reset()

tmap_mode("plot")

tm_shape(vegetationshoehe_LFI) +
  tm_raster(
    title = "Vegetation height"
  )

tm_shape(vegetationshoehe_LFI) +
  tm_raster(
    title = "Vegetation height"
  ) +
  tm_shape(wildschwein_BE) +
  tm_dots(size = 0.02)
```

## Task 5

```{r}

vegetation_values <- terra::extract(
  vegetationshoehe_LFI,
  terra::vect(wildschwein_BE)
)

head(vegetation_values)


wildschwein_BE_veg <- wildschwein_BE |>
  mutate(
    vegetationshoehe_LFI = vegetation_values[[2]]
  ) |>
  select(TierID, TierName, DatetimeUTC, vegetationshoehe_LFI, geometry)

wildschwein_BE_veg


summary(wildschwein_BE_veg$vegetationshoehe_LFI)

wildschwein_BE_veg |>
  st_drop_geometry() |>
  group_by(TierName) |>
  summarise(
    n = n(),
    mean_vegetation_height = mean(vegetationshoehe_LFI, na.rm = TRUE),
    median_vegetation_height = median(vegetationshoehe_LFI, na.rm = TRUE),
    min_vegetation_height = min(vegetationshoehe_LFI, na.rm = TRUE),
    max_vegetation_height = max(vegetationshoehe_LFI, na.rm = TRUE)
  )


wildschwein_BE_veg <- wildschwein_BE_veg |>
  mutate(
    DatetimeUTC = ymd_hms(DatetimeUTC),
    hour = hour(DatetimeUTC)
  )


ggplot(wildschwein_BE_veg, aes(x = hour, y = vegetationshoehe_LFI)) +
  geom_boxplot(aes(group = hour)) +
  scale_x_continuous(breaks = 0:23) +
  labs(
    title = "Vegetation height at wild boar locations by hour",
    x = "Hour of day",
    y = "Vegetation height"
  ) +
  theme_minimal()


ggplot(wildschwein_BE_veg, aes(x = TierName, y = vegetationshoehe_LFI)) +
  geom_boxplot() +
  labs(
    title = "Vegetation height at wild boar locations by individual",
    x = "Wild boar",
    y = "Vegetation height"
  ) +
  theme_minimal()
```
